---
title: "Milestone #3"
output: html_document
---
## Background
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE)
library(tidyverse)
library(readxl)
library(janitor)
```
The purpose of this project is to analyze the differences in educational outcome between students in different educational circumstances across the country. I plan to include data from as many of the 50 states as I can access, though some states are much more restrictive than others. Below, I have selected some data that was readily available to me on the Massachusetts Department of Education Website. Included is data detailing SAT performance by school district, per pupil expenditures by school district, and staff demographics by school district. 
```{r}
MA_SAT_18_19 <- read_excel("sat_performance.xlsx", 
                        col_types = c("text", "numeric",
                                      "numeric", "numeric",
                                      "logical", "numeric"),
                        skip = 1) %>% 
  clean_names() 

MA_PPE_18 <- read_excel("PerPupilExpenditures.xlsx",
                        col_types = c("text", "numeric",
                                      "numeric", "text",
                                      "text", "text",
                                      "text", "text",
                                      "logical"),
                        skip = 1) %>% 
  clean_names() %>% 
  select(-x9)

mass_staff_demographics <- read_excel("staffracegender.xlsx",
                                      skip = 1)
```
I then joined the SAT scores data with the per pupil expenditure data, which will allow me to sort by level of expenditure per student and draw conclusions from the district's average SAT score. The SAT scores data includes another breakdown; one for science/math scores and one for reading/writing.
```{r}
mass_joined <- MA_SAT_18_19 %>% 
  full_join(MA_PPE_18, by = c("district_name" = "district_name",
                              "district_code" = "district_code"))
```
I plan to draw conclusions across the different sets of data across all of the states, and look at how differences in demographic composition, per pupil expenditure, school district salary, and district salary influence outcomes as measured by ACT/SAT score and college admission statistics. 
```{r}
mass_joined %>% 
  arrange(desc(in_district_expenditures)) %>% 
  View()
```

